Reanalysis Step 7: Missingness Pattern Analysis

### 2025-09-01 reanalysis-7-missingness_patterns.qmd
fix ai idiocy

# Process each variable in the table
for (var in available_vars) {
  if (var %in% names(original_data)) {
    if (var %in% factor_vars && is.factor(original_data[[var]])) {
      # For factor variables, calculate SMD for each level
      var_levels <- levels(original_data[[var]])
      for (level in var_levels) {
        # Calculate proportion in each group
        prop_excluded <- mean(original_data[[var]][original_data$excluded] == level, na.rm = TRUE)
        prop_included <- mean(original_data[[var]][!original_data$excluded] == level, na.rm = TRUE)

@@

        smd_data <- rbind(smd_data, data.frame(
          Variable = paste0(var, ": ", level),
          SMD = smd_val,
          VarType = "Factor",
          BaseVar = var,
          stringsAsFactors = FALSE
        ))
      }
    } else {
      # For continuous variables, calculate single SMD
      if (sum(!is.na(original_data[[var]])) > 0) {
        mean_excluded <- mean(original_data[[var]][original_data$excluded], na.rm = TRUE)
        mean_included <- mean(original_data[[var]][!original_data$excluded], na.rm = TRUE)
        sd_pooled <- sqrt((var(original_data[[var]][original_data$excluded], na.rm = TRUE) + 
                          var(original_data[[var]][!original_data$excluded], na.rm = TRUE)) / 2)

        smd_val <- if (!is.na(sd_pooled) && sd_pooled > 0) {
          abs(mean_excluded - mean_included) / sd_pooled
        } else {
          0
        }

>>>>

# Process each variable in the table
for (var in available_vars) {
  if (var %in% names(original_data)) {
    # Check if it's a factor or character that should be treated as factor
    is_factor_var <- (var %in% factor_vars) || is.factor(original_data[[var]]) || is.character(original_data[[var]])

    if (is_factor_var) {
      # For factor/categorical variables, calculate SMD for each level
      # Convert to factor if needed
      if (!is.factor(original_data[[var]])) {
        original_data[[var]] <- as.factor(original_data[[var]])
      }

      var_levels <- levels(original_data[[var]])
      for (level in var_levels) {
        # Calculate proportion in each group
        prop_excluded <- mean(original_data[[var]][original_data$excluded] == level, na.rm = TRUE)
        prop_included <- mean(original_data[[var]][!original_data$excluded] == level, na.rm = TRUE)

@@

        smd_data <- rbind(smd_data, data.frame(
          Variable = paste0(var, ": ", level),
          SMD = smd_val,
          VarType = "Factor",
          BaseVar = var,
          stringsAsFactors = FALSE
        ))
      }
    } else if (is.numeric(original_data[[var]])) {
      # For continuous variables, calculate single SMD
      if (sum(!is.na(original_data[[var]])) > 0) {
        # Calculate means
        mean_excluded <- mean(original_data[[var]][original_data$excluded], na.rm = TRUE)
        mean_included <- mean(original_data[[var]][!original_data$excluded], na.rm = TRUE)

        # Calculate pooled standard deviation
        var_excluded <- var(original_data[[var]][original_data$excluded], na.rm = TRUE)
        var_included <- var(original_data[[var]][!original_data$excluded], na.rm = TRUE)

        # Check if we have valid variances
        if (!is.na(var_excluded) && !is.na(var_included)) {
          sd_pooled <- sqrt((var_excluded + var_included) / 2)

          smd_val <- if (!is.na(sd_pooled) && sd_pooled > 0) {
            abs(mean_excluded - mean_included) / sd_pooled
          } else {
            0
          }
        } else {
          smd_val <- 0
        }
